Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-08-09 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
Appendices figures/tables notations corrected

\renewcommand\thefigure{A.\arabic{figure}}
\renewcommand\thetable{A.\arabic{table}}

@@

\subsection{Effect of the starting tree on Bayesian inference}
We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference for decreasing the computing time.
To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
We tested this effect on five trees with the five levels of missing data (i.e. first tree: $M_L$=0\%, $M_F$=0\% and $N_C$=100 (i.e. 0\% ``missing''); second tree: $M_L$=10\%, $M_F$=10\% and $N_C$=90, etc.) on the first 20 simulation chains.
We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
We then compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
We found no significant effect of using the ``true'' or a random tree as a starting tree on the ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).

\begin{table}[ht]
\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.}
\label{Tab_lm_res}
\centering
\begin{tabular}{rlcccc}
  \hline
 & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
RF & (Intercept) & 0.8100 & 0.0122 & 66.16 & \textbf{0.0000} \\ 
   & starting:true & 0.0019 & 0.0173 & 0.11 & 0.9125 \\
Tr & (Intercept) & 0.4424 & 0.0418 & 10.59 & \textbf{0.0000} \\ 
   & starting:true & 0.0159 & 0.0591 & 0.27 & 0.7887 \\ 
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[keepaspectratio=true, width=\textwidth]{SupplementaryFigures/Starting_tree.pdf}
\caption{Effect of using the ``true'' tree (black) or a random tree (red) as the starting tree for the Bayesian inference. The x axis, represent the levels of missing data ()}
\label{Fig_startingTree}
\end{figure}

\subsection{Morphological rates estimations results}
To test the sensitivity of the Bayesian prior on the morphological rates distribution shape ($\alpha$ = 0.5), we compared the shape of the distribution of our generated morphological characters ($\alpha_{TRUE}$) to the shape of the distribution of morphological characters rate estimated by the Bayesian inference ($\alpha_{EST}$).
The $\alpha_{TRUE}$ parameter was estimated by fitting a gamma distribution to the 100 individual morphological rates (randomly sampled from a gamma distribution with $\alpha$ = 0.5) using the \texttt{fitdistr} function from the \texttt{MASS R} package \citep{MASS}.
We measured the difference as:
\begin{equation}
\Delta\alpha=\alpha_{EST} - \alpha_{TRUE}
\end{equation}
We then calculated the mode and the 95\% and 50\% confidence intervals over the 50 chains using the \texttt{hdr} function from the \texttt{hdrcde R} package \citep{hdrcde}.
The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Rates_estimates.R}.

>>>>

\renewcommand\thefigure{A\arabic{figure}}
\renewcommand\thetable{A\arabic{table}}

@@

\subsection{Effect of the starting tree on Bayesian inference}
We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference to decrease the computing time needed.
To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random starting tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
We tested this effect on five trees with the five levels of missing data (i.e. first tree: $M_L$=0\%, $M_F$=0\% and $N_C$=100 (i.e. 0\% ``missing''); second tree: $M_L$=10\%, $M_F$=10\% and $N_C$=90, etc.) on the first 20 simulation chains.
We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
We also compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
% NC: Why the regressions? What are they telling you? Why simple and nested?
The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
We found no significant effect of using the ``true'' instead of a random tree as a starting tree on our ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).

\begin{table}[ht]
\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.} % NC: I'm not sure what these two tables are telling you. Clarify. The caption says nothing about the random starting trees.
\label{Tab_lm_res}
\centering
\begin{tabular}{rlcccc}
  \hline
 & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
RF & (Intercept) & 0.8100 & 0.0122 & 66.16 & \textbf{0.0000} \\ 
   & starting:true & 0.0019 & 0.0173 & 0.11 & 0.9125 \\
Tr & (Intercept) & 0.4424 & 0.0418 & 10.59 & \textbf{0.0000} \\ 
   & starting:true & 0.0159 & 0.0591 & 0.27 & 0.7887 \\ 
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[keepaspectratio=true, width=\textwidth]{SupplementaryFigures/Starting_tree.pdf}
\caption{Effect of using the ``true'' tree (black) or a random tree (red) as the starting tree for the Bayesian inference. The x axis, represents the amount of missing data ()} % NC: What's missing from the brackets here or are they a typo?
\label{Fig_startingTree}
\end{figure}

\subsection{Morphological rates estimates results}
To test the sensitivity of the Bayesian prior on the morphological rates distribution shape ($\alpha$ = 0.5), we compared the shape of the distribution of our generated morphological characters ($\alpha_{TRUE}$) to the shape of the distribution of morphological character rates estimated by the Bayesian inference ($\alpha_{EST}$).
The $\alpha_{TRUE}$ parameter was estimated by fitting a gamma distribution to the 100 individual morphological rates (randomly sampled from a gamma distribution with $\alpha$ = 0.5) using the \texttt{fitdistr} function from the \texttt{MASS} R package \citep{MASS}.
We measured the difference as:
\begin{equation}
\Delta\alpha=\alpha_{EST} - \alpha_{TRUE}
\end{equation}
We then calculated the mode and the 95\% and 50\% confidence intervals over the 50 chains using the \texttt{hdr} function from the \texttt{hdrcde} R package \citep{hdrcde}.
The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Rates_estimates.R}.

### 2015-08-09 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
Added appendix tables and giros prefix
Appendices figures/tables notations corrected

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
%\setcounter{secnumdepth}{0} 
\begin{document}

@@

>>>>

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}

\renewcommand\thefigure{B\arabic{figure}}
\renewcommand\thetable{B\arabic{table}}

%\setcounter{secnumdepth}{0} 
\begin{document}

@@

\subsection{Differences between the ``true'' and the ``best'' trees.}
% NC: Need a line or two to explain the motivation of this analysis. Why did you do it? (not just because reviewers asked you to!)
We compared our ``true'' trees (i.e. the trees \textbf{used to create} the ``complete'' matrices) to the ``best'' trees (i.e. the trees \textbf{inferred from} % NC: I thought we weren't using the word inferred
 the ``complete'' matrices) in figure ~\ref{Fig_Supp_True_Best}.
Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer % NC: again
phylogenies from this particular matrices (i.e. small sized and generated using specific algorithms).
This does not affect, however, the results of our analysis since we deliberately compared the the ``missing-data'' trees to the ``best'' tree rather than to the ``true'' tree.

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryFigures/True_vs_Best_trees.pdf}
    \caption{Pairwise comparisons among the 50 ``true'' trees and the 50 ``best'' trees from the Maximum Likelihood and Bayesian inference methods. The horizontal blue and red lines represent, respectively, the 95\% and 50\% confidence intervals.} 
\label{Fig_Supp_True_Best} 
\end{figure}

### 2015-08-09 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Appendices figures/tables notations corrected

\renewcommand\thefigure{C.\arabic{figure}}
\renewcommand\thetable{C.\arabic{table}}

@@

The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper. For clarity, in the paper, we focused on the results of the effects of our missing data parameters on the Maximum Likelihood trees topology and the Bayesian consensus trees topology. The following additional results presented here give a greater insight into the effect of our missing data parameters on the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies. Also we present here the pairwise comparisons between each parameters states for the Maximum Likelihood tree topology, the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.

>>>>

\renewcommand\thefigure{C\arabic{figure}}
\renewcommand\thetable{C\arabic{table}}

@@

The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper. For clarity, in the paper, we focused on the results of the effects of our missing data parameters on the Maximum Likelihood trees topology and the Bayesian consensus trees topology. The following additional results presented here give a greater insight into the effect of our missing data parameters on the Maximum Likelihood bootstrapped tree topologies and the Bayesian posterior tree distribution topologies. Also we present here the pairwise comparisons between each parameters states for the Maximum Likelihood tree topology, the Maximum Likelihood bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
